### 2016-11-25 fhe.tex

\begin{section}{Fully homomorphic encryption scheme~\cite{DBLP:conf/crypto/GentrySW13}}

  In this section, we will instantiate the GSW fully homomorphic encryption scheme.

  As in the original paper, we will firstly give a basic encryption scheme.

  \begin{desciption}
  \item[$\Setup(1^\lambda)$]: 
    \begin{enumerate}
    \item We
    \end{enumerate}
  \end{description}

\end{section}