### 2016-12-28 LTF.tex
modif 28/12/2016 day

Lossy trapdoor functions are proved to be very usefull for the construction of CCA-secure encryption scheme. 
There already exists many efficient constructions of all-but-one lossy trapdoor functions with various assumptions~\cite{DBLP:journals/siamcomp/PeikertW11, DBLP:conf/crypto/BoldyrevaFO08}

\begin{section}{All-but-one lossy trapdoor function}

\end{section}