### 2016-12-28 lemma.tex

\begin{section}{Polynomial theorems}

  \begin{lemma}~\cite{DBLP:conf/asiacrypt/Lyubashevsky16}
    For any polynomial $a \in \mathbb{Z}[x]$ of degree $n$ and a polynomial $b \in \mathbb{Z}[x]$ of degree $m$ each of whose coefficient is chosen at random modulo $q$.
    Let $c$ be the product of the two polynomials $c = a \cdot b$.
    \begin{align*}
    c &= c_0 + c_1 x + \cdots + c_n x^n + \cdots + c_{m+n}x^{m+n}
    \end{align*}

    The coefficients $c_n, \dots, c_{m+n}$ are jointly univormly random modulo q \wrt the polynomial $a$.
  \end{lemma}

\end{section}