### 2017-03-03 draftNEFHE.tex

\begin{section}{Non-error Somewhat Fully Homomorphic encryption scheme}
  Here we firstly give the blueprint and the main idea of our construction, This idea is initialized by the work of Fully homomorphic encryption scheme for $NC^1$, which give us a fully homomorphic encryption scheme without error, but using only an mod $2$ additively homomorphic encryption scheme. However, in their work the ciphertext size is doubling when we perform a homomorphic operation on the ciphertext.
  \begin{itemize}
  \item 
  \end{itemize}
\end{section}